---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# jovanovic2025

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/jovanovic2025)](https://CRAN.R-project.org/package=jovanovic2025)
[![R-CMD-check](https://github.com/pandionlabs/jovanovic2025/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/pandionlabs/jovanovic2025/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This is the companion code and data to Jovanovic et al. 2025. 

## Installation

You can install the development version of jovanovic2025 from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("pandionlabs/jovanovic2025")
```

## Example

Load the package like this:

```{r example}
library(jovanovic2025)
```

```{r loaddata}
# Import the data from a tab delimited ascii file.
# MasterThesisData <- read.table(
#   file = "data/raw/MasterThesisData2024.csv",
#   header = TRUE,
#   sep = ",",
#   na.strings = "NA",
#   stringsAsFactors = TRUE,
#   dec = "."
# )

# Import the data with MasterThesisData2024
TreMs <- clean_data(MasterThesisData2024) 
```


```{r}

TreeIdentitiesSummary <- summarize_identities(
  TreMs, 
  grouping_variable = TreeIdentities2
  )

  # write.table(TreeIdentitiesSummary, file = "data/derivatives/TreeIdentitiesSummary.csv", sep = ",", quote = FALSE, row.names = FALSE)


DeadwoodIdentitiesGroupedSummary <- summarize_identities(
  TreMs, 
  grouping_variable = DeadwoodIdentitiesGrouped
  )

# write.table(DeadwoodIdentitiesGroupedSummary, file = "data/derivatives/DeadwoodIdentitiesGroupedSummary.csv", sep = ",", quote = FALSE, row.names = FALSE)
```

```{r}
#| echo: false
#| label: tbl-TreeIdentitiesSummary
#| tbl-cap: "TreeIdentitiesSummary"
knitr::kable(TreeIdentitiesSummary)
```

```{r}
#| echo: false
#| label: tbl-DeadwoodIdentitiesGroupedSummary
#| tbl-cap: "DeadwoodIdentitiesGroupedSummary"
knitr::kable(DeadwoodIdentitiesGroupedSummary)
```

## Make a count table by decay stage


```{r}
count_table <- make_count_table_by_decay(
  TreMs, 
  save_name = "data/derivatives/Count_Table_by_Decay_Stage.csv"
)
```

## Create models

In the create models section, we build a series of models associating various predictor (y) variables with three varialbes: `GroupedTreeSpecies, Treedata.DBH_cm, and Treedata.Tree_Decay`. The modelling family for each model is either poisson or glmmTMB::nbinom2.

The `get_model_cols()` function creates a tibble that specifies each predictor variable and the model family used in each model


```{r}
model_parameters <- get_model_cols()
```


```{r}
#| echo: false
model_parameters_tb <- 
  tibble::tribble(
    ~variable,                ~model_family,
    "DecomposedCrack",           "poisson",
    "WoodpeckerCavities",        "poisson",
    "Concavities",               "poisson",
    "WoodpeckerConcavities",     "poisson",
    "Rotholes",                  "poisson",
    "InsectGalleries",           "poisson",
    "ExposedSapwood",            "glmmTMB::nbinom2",
    "ExposedHeartwood",          "glmmTMB::nbinom2",
    "ExposedSapwoodHeartwood",   "glmmTMB::nbinom2",
    "PerennialFungi",            "glmmTMB::nbinom2",
    "Ephermalfungi",             "glmmTMB::nbinom2",
    "EphrmalPerennialFungi",     "glmmTMB::nbinom2",
    "Epiphytes",                 "glmmTMB::nbinom2",
    "DeadwooodShelter",          "glmmTMB::nbinom2",
    "StumpStructures",           "poisson",
    "DeadwooodShelterStumpStructures", "glmmTMB::nbinom2",
    "LogStructures",             "glmmTMB::nbinom2",
    "WoodyDebris",               "glmmTMB::nbinom2",
    "ExposedRoots",              "glmmTMB::nbinom2",
    "Abundance",                 "glmmTMB::nbinom2",
    "Richness",                  "glmmTMB::nbinom2"
  )
knitr::kable(model_parameters_tb)
```

The models can be produced with `create_models()` which will create each of the models specified in `get_model_cols()` and then save the model summary, residuals, and results from three tests of outliers, dispersion, and zero inflation. All these models are results are returned as a dataframe.

```{r}
#| echo: true
#| warning: false
#| output: false
#| message: false
mod_df <- create_models(TreMs)
```