---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# jovanovic2025

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/jovanovic2025)](https://CRAN.R-project.org/package=jovanovic2025)
<!-- badges: end -->

This is the companion code and data to Jovanovic et al. 2025. 

## Installation

You can install the development version of jovanovic2025 from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("pandionlabs/jovanovic_2025")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(jovanovic2025)
## basic example code
```

```{r loaddata}
# Import the data from a tab delimited ascii file.
MasterThesisData <- read.table(
  file = "data/raw/MasterThesisData2024.csv",
  header = TRUE,
  sep = ",",
  na.strings = "NA",
  stringsAsFactors = TRUE,
  dec = "."
)

TreMs <- clean_data(MasterThesisData)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
