---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# jovanovic2025

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/jovanovic2025)](https://CRAN.R-project.org/package=jovanovic2025)
[![R-CMD-check](https://github.com/pandionlabs/jovanovic2025/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/pandionlabs/jovanovic2025/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This is the companion code and data to Jovanovic et al. 2025. 

## Installation

You can install the development version of jovanovic2025 from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("pandionlabs/jovanovic2025")
```

## Example

Load the package like this:

```{r example}
library(jovanovic2025)
```

```{r loaddata}
# Import the data from a tab delimited ascii file.
# MasterThesisData <- read.table(
#   file = "data/raw/MasterThesisData2024.csv",
#   header = TRUE,
#   sep = ",",
#   na.strings = "NA",
#   stringsAsFactors = TRUE,
#   dec = "."
# )

# Import the data with MasterThesisData2024
TreMs <- clean_data(MasterThesisData2024) 
```


```{r}

TreeIdentitiesSummary <- summarize_identities(
  TreMs, 
  grouping_variable = TreeIdentities2
  )

  # write.table(TreeIdentitiesSummary, file = "data/derivatives/TreeIdentitiesSummary.csv", sep = ",", quote = FALSE, row.names = FALSE)


DeadwoodIdentitiesGroupedSummary <- summarize_identities(
  TreMs, 
  grouping_variable = DeadwoodIdentitiesGrouped
  )

# write.table(DeadwoodIdentitiesGroupedSummary, file = "data/derivatives/DeadwoodIdentitiesGroupedSummary.csv", sep = ",", quote = FALSE, row.names = FALSE)
```

```{r}
#| echo: false
#| label: tbl-TreeIdentitiesSummary
#| tbl-cap: "TreeIdentitiesSummary"
knitr::kable(TreeIdentitiesSummary)
```

```{r}
#| echo: false
#| label: tbl-DeadwoodIdentitiesGroupedSummary
#| tbl-cap: "DeadwoodIdentitiesGroupedSummary"
knitr::kable(DeadwoodIdentitiesGroupedSummary)
```

## Make a count table by decay stage


```{r}
count_table <- make_count_table_by_decay(
  TreMs, 
  save_name = "data/derivatives/Count_Table_by_Decay_Stage.csv"
)
```

## Create models

In the create models section, we build a series of models associating various predictor (y) variables with three varialbes: `GroupedTreeSpecies, Treedata.DBH_cm, and Treedata.Tree_Decay`. The modelling family for each model is either poisson or glmmTMB::nbinom2.

The `get_model_cols()` function creates a tibble that specifies each predictor variable and the model family used in each model


```{r}
model_parameters <- get_model_cols()
```


```{r}
#| echo: false
model_parameters_tb <- 
  tibble::tribble(
    ~variable,                ~model_family,
    "DecomposedCrack",           "poisson",
    "WoodpeckerCavities",        "poisson",
    "Concavities",               "poisson",
    "WoodpeckerConcavities",     "poisson",
    "Rotholes",                  "poisson",
    "InsectGalleries",           "poisson",
    "ExposedSapwood",            "poisson",
    "ExposedHeartwood",          "poisson",
    "ExposedSapwoodHeartwood",   "poisson",
    "PerennialFungi",            "poisson",
    "Ephermalfungi",             "poisson",
    "EphrmalPerennialFungi",     "poisson",
    "Epiphytes",                 "poisson",
    "DeadwooodShelter",          "poisson",
    "StumpStructures",           "poisson",
    "DeadwooodShelterStumpStructures", "poisson",
    "LogStructures",             "poisson",
    "WoodyDebris",               "poisson",
    "ExposedRoots",              "poisson",
    "Abundance",                 "poisson",
    "Richness",                  "poisson",
    "DecomposedCrack",           "glmmTMB::nbinom2",
    "WoodpeckerCavities",        "glmmTMB::nbinom2",
    "Concavities",               "glmmTMB::nbinom2",
    "WoodpeckerConcavities",     "glmmTMB::nbinom2",
    "Rotholes",                  "glmmTMB::nbinom2",
    "InsectGalleries",           "glmmTMB::nbinom2",
    "ExposedSapwood",            "glmmTMB::nbinom2",
    "ExposedHeartwood",          "glmmTMB::nbinom2",
    "ExposedSapwoodHeartwood",   "glmmTMB::nbinom2",
    "PerennialFungi",            "glmmTMB::nbinom2",
    "Ephermalfungi",             "glmmTMB::nbinom2",
    "EphrmalPerennialFungi",     "glmmTMB::nbinom2",
    "Epiphytes",                 "glmmTMB::nbinom2",
    "DeadwoodShelter",          "glmmTMB::nbinom2",
    "StumpStructures",           "glmmTMB::nbinom2",
    "DeadwooodShelterStumpStructures", "glmmTMB::nbinom2",
    "LogStructures",             "glmmTMB::nbinom2",
    "WoodyDebris",               "glmmTMB::nbinom2",
    "ExposedRoots",              "glmmTMB::nbinom2",
    "Abundance",                 "glmmTMB::nbinom2",
    "Richness",                  "glmmTMB::nbinom2"
  ) |> 
    dplyr::mutate(
      model_formula = glue::glue("{variable} ~ GroupedTreeSpecies + Treedata.DBH_cm + Treedata.Tree_Decay + (1\\|Plot)")
      )

knitr::kable(model_parameters_tb)
```

The models can be produced with `create_models()` which will create each of the models specified in `get_model_cols()` and then save the model summary, residuals, and results from three tests of outliers, dispersion, and zero inflation. All these models are results are returned as a dataframe.

```{r}
#| echo: true
#| warning: false
#| output: false
#| message: false
mod_df <- create_models(TreMs)
```

Access results of those models with the following

```{r}
get_formula(mod_df, 1)
get_residuals(mod_df, 1)
get_test_outliers(mod_df, 1)
get_test_dispersion(mod_df, 1)
get_test_zero_inflation(mod_df, 1)
```

## Get models

We can delve into the model dataframe (`mod_df`) with `get_model_object`
```{r}
# Get a model object by index
get_model_object(mod_df, 2)

# Or get a model object using filter 

mod_df |> 
  dplyr::filter(
    variable == "Rotholes",
    model_family == "poisson"
    ) |> 
  get_model_object()
```

## Make a graph

Use ggeffects to predict on a model and then graph it with ggplot.

```{r}
Predicted_RichnessStump <- 
mod_df |> 
  dplyr::filter(
    variable == "Richness",
    model_family == "nbinom2"
    ) |> 
  get_model_object() |> 
  ggeffects::ggpredict(
    terms = c("Treedata.DBH_cm", "Treedata.Tree_Decay[Decay stage 1, Decay stage 5]"),
    interval = "confidence",
    full.data = TRUE,
    bias_correction = TRUE,
    ci_level = 0.95
  )

library(ggplot2)
ggplot() +
  geom_smooth(
    data = Predicted_RichnessStump, 
    mapping = aes(x = x, y = predicted, colour = group)
    ) +
  geom_rug(
    data = TreMs, 
    mapping = aes(x = Treedata.DBH_cm, y = Richness), 
    col = "steelblue",
    alpha=0.1, 
    size=1
    ) +
  xlab("Diameter (cm)") + ylab("TreM Richness on Stumps")+ 
  scale_x_continuous(n.breaks = 5) +
  theme_bw() +
  theme(
    text = element_text(size=15), 
    legend.title = element_blank(),
    plot.margin = margin(0.5, 2, 0.5, 0.5),
    panel.border = element_blank(), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    axis.line = element_line(colour = "black")
    ) +
  scale_linetype(guide = "none") +
  scale_color_manual(values = c("Decay stage 1" = "#849324", "Decay stage 5" = "#f26419")) +
  scale_fill_manual(values = c("Decay stage 1" = "#849324", "Decay stage 5" = "#f26419"))
```